Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-03-28 ident_manu.Rnw
better draft done

\begin{abstract}
\noindent This study addresses quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  A systematic framework is used to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates and less sensitive to physical or chemical parameters.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.
\end{abstract}

@@

The development and application of a model represents a tradeoff between characteristics expected from the output or provided by the structural components. An idealized model is sufficiently generalizable across systems, provides results that are precise given the inputs, and includes components that are realistic descriptions of actual processes \citep{Levins66}. Given that these characteristics cannot be simultaneously achieved, models are developed in partial dependence of reality and theoretical constructs, completely separate from both, or dependent on one or the other \citep{Morrison99,Ganju16}.  These challenges are analagous to the well-known bias-variance tradeoff in statistical models that balances the competing objectives of over- and under-fitting to an observed dataset. Process-based models are more commonly imbalanced between reality and theory, such that most are over-parameterized in an attempt to completely describe reality \citep{Denman03,Nossent12,Petrucci14}.  Quantitative limitations of over-parameterization are analagous to degrees of freedom in standard statistical models as free parameters cannot be numerically estimated when constrained to an observed dataset \citep{Kirchner06}.  More importantly, over-parameterization can limit use across systems outside of the data domain and impose uncertainty in model predictions as realistic values for every variable may not be known or inaccurately applied from existing studies \citep{Durand02,Refsgaard07,Wade08}.

@@

This study describes a parameter sensitivity and identifiability analysis of a \ac{zerod} unit of a larger spatial-temporal model of hypoxia dynamics on the \ac{lcs}.  The objectives were to provide a statistical approach that demonstrates numerical limitations of parameter sets for model calibration and provide a framework for selecting parameters within the identifability constraints.  The specific goals were to \begin{inparaenum}[1\upshape)]
\item identify the parameters that have the greatest influence on state variables using local sensitivity analysis,
\item quantify the identifiability of subsets of the total parameter space based on sensitivity,
\item and provide a set of heuristics for choosing parameters based on sensitivity, identifiability, and parameter categories.
\end{inparaenum}
In addition to \ac{do}, other state variables that were evaluated included ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus. In general, we provide empirical results to support the assumption that models are generally over-parameterized and only a finite and smaller subset of the complete parameter set can be optimized. Although concerns of identifiability constraints have been shown in the literature, rulesets are not readily available that define how parameters should be selected given tradeoffs betwen precision, sensitivity, and identifiability, especially for biogeochemical models of ecosystem processes.

@@

Hypoxic events, defined  as $<$2 mg L$^{-1}$ of \ac{do} ($<$ 64 mmol m$^{-3}$), occur seasonally in bottom waters in the northern \ac{gom}.  The hypoxic area averages 15,540 km$^2$ annually (1993-2015) with minimum concentrations observed from late spring to early fall.  Seasonal variation is strongly related to carbon and nutrient export from the \ac{marb} \citep{Lohrenz08,Bianchi10}, whereas hydrologic variation, currents, and wind patterns can affect vertical salinity gradients that contribute to the formation of hypoxia \citep{Wiseman97,Paerl98,Obenour15}. The \ac{cgem} was developed to describe hypoxia dynamics on the \ac{lcs} and includes elements from the Navy Coastal Ocean Model \citep{Martin00} for hydrodynamics and a biogeochemical model with multiple plankton groups, water-column metabolism, and sediment diagenesis \citep{Eldridge10}.  The hydrodynamic component of \ac{cgem} provides a spatially-explicit description of hypoxia using an orthogonal grid with an approximate horizontal resolution of 1.9 km$^2$ and twenty equally-spaced vertical sigma layers on the shelf.  The biogeochemical component includes equations for 36 state variables including six phytoplankton groups (with nitrogen and phosophorus quotas for each), two zooplankton groups, nitrate, ammonium, phosphate, dissolved inorganic carbon, oxygen, silica, and multiple variables for dissolved and particulate organic matter from different sources.

@@

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included \Sexpr{length(p1z1pars)} parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ \Sexpr{nparms$Optics} parameters), organic matter (\Sexpr{nparms$`Organic Matter`}), phytoplankton (\Sexpr{nparms$Phytoplankton}), temperature (\Sexpr{nparms$Temperature}), and zooplankton (\Sexpr{nparms$Zooplankton}).  A full description of the model parameters is available as an appendix in \citetalias{LehrterIP}.

@@

A local sensitivity analysis was performed by evaluating the change in state variables following perturbation of each parameter from its original value \citep{Soetaert10,RDCT17}. Parameters were individually perturbed by 50\% of the original values and sensitivity $S$ was estimated for each time step $i$ given a change for parameter $j$ as:

@@

Parameter selection for model calibration must consider the competing objectives of increased precision with parameter inclusion and reduced identifability as it relates to optimization.  An additional challenge is a large number of combinations of parameter sets, which complicates selection given sensitivity differences and desired ecological categories of each parameter (e.g., parameters for a group of related structural equations could be of interest).  \cref{fig:combnex} provides a simple graphic of the unique number of combinations that are possible for different subsets of `complete' parameter sets of different sizes (i.e., $n$ choose $k$ combinations, $n!/\left(k!\left(n-k\right)!\right)$).  The number of unique combinations increases with the total parameters in the set and is also maximized for moderate selections (e.g., selecting half the total).  For example, over 10$^{14}$ combinations are possible by selecting 25 parameters from a set of 50.

A set of heuristics was developed that addresses the tradeoff in model complexity and identifiability given the challenges described above \citep[see also][]{Wagener01}.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability between parameters).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences by model users given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally.
\end{inparaenum} The selection rules produced seven subsets of parameters that could further be used to optimize model calibration.

\subsection{Model calibration and extrapolation}

@@

Each parameter subset was calibrated to most closely match the model output to the observed time series.  Parameter values were searched using the optim function in R that implemented the limited memory modification of the BFGS quasi-Newton method \citep{Byrd95,Nocedal06,RDCT17}.  This opimization scheme searches for values within the user-defined lower and upper bounds of each parameter.  We set the algorithm to search within +/- 50\% of the default parameter values.  While we recognize that `actual' parameter values could possibly be outside of this range, constraining the values to those tested herein was expected to produce output that was more interpretable within the context of the sensitivity functions defined above.  Moreover, precise limits on each parameter were less of a concern for the current analysis as our primary goal was to demonstrate differences in precision as a function of sensitivity and identifiability.  The optimal values for each parameter were identified based on convergence of the model results as a minimization of the \ac{rmse} between the observed and modelled data.

As a final analysis, effects of parameter changes on the larger 3-dimensional model were evaluated to demonstrate that results from the \ac{zerod} FishTank model can be extrapolated.  For simplicity, only the most sensitive parameter in each category was selected.  The \ac{cgem} model was run for one year using the default parameter conditions and by increasing each of the selected parameters by 50\%.  This produced six spatially-referenced time series from which the changes in the daily hypoxic area (number of cells $<$ 64 mmol \ac{do}) were estimated in the nothern \ac{gom}.

@@

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$.  Parameter identifiability for \ac{do} decreased (increasing $\gamma$) at different rates with increasing size of parameter subsets depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets in the phytoplankton (\Sexpr{catpersO2$Phytoplankton[1]}\% of subsets less than $\gamma = 15$, \Sexpr{catpersO2$Phytoplankton[2]}\% less than $\gamma = 10$) and zooplankton categories (\Sexpr{catpersO2$Zooplankton[1]}\% less than $\gamma = 15$, \Sexpr{catpersO2$Zooplankton[2]}\% less than $\gamma = 10$), whereas all combinations were identifiable for optics (\Sexpr{catpersO2$Optics[1]}\% less than $\gamma = 15, 10$) and a majority identifiable for organic matter (\Sexpr{catpersO2$`Organic Matter`[1]}\% less than $\gamma = 15$, \Sexpr{catpersO2$`Organic Matter`[2]}\% less than $\gamma = 10$). Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifability was \Sexpr{toppersO2$top1[1]}\% less than $\gamma = 15, 10$ for the top parameter in each category, \Sexpr{toppersO2$top2[1]}\% and \Sexpr{toppersO2$top2[2]}\% for the top two, \Sexpr{toppersO2$top3[1]}\% and \Sexpr{toppersO2$top3[2]}\% for the top three, and \Sexpr{toppersO2$top4[1]}\% and \Sexpr{toppersO2$top4[2]}\% for the top four.  Results for the remaining state variables had similar patterns in identifiability with increasing size of parameter subsets and selection categories, although differences in identifiabiltiy between state variables was observed (\cref{fig:identploall}).  Most notably, nitrate was consistently the least identifiable variable (highest overall $\gamma$), whereas \ac{do} was most identifiable.

@@

\subsection{Model calibration and extrapolation}

@@

Calibration of parameter values for each of the seven selection heuristics produced model output that varied in precision and computational effort to identify optimal values (\cref{tab:calib}).  The seven evaluated subsets were those for \ac{do} in each of the four categories of parameters in \cref{tab:heurist1}, including the single parameter for temperature.  The subsets in \cref{tab:heurist2,tab:heurist3} for the parameters selected independent of parameter category and equally within each category were also evaluated.  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only \Sexpr{toeng(temp$its)[[1]]} iterations were required to optimize the single parameter selected for the temperature category, whereas \Sexpr{top$its} iterations were required for the \Sexpr{toeng(top$nprm)[[1]]} parameters selected independent of category.  A slight positive assocation between collinearity ($\gamma$) and number of iterations was also observed, such that computational effort  to identify optimal values for a parameter set increased with lower identifiability (higher $\gamma$).  Of the seven parameter subsets, selection independent of category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} \Sexpr{strt}, reduction to \Sexpr{scinot(top$rmse, digits = 3)}, \Sexpr{scinot(cattop$rmse, digits = 3)}, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction (\Sexpr{scinot(optics$rmse, digits = 3)}).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 (\Sexpr{scinot(optics$avee)}, \Sexpr{toeng(optics$nprm)} parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 (\Sexpr{scinot(top$avee)}, \Sexpr{toeng(top$nprm)} parameters, \cref{tab:calib}).

@@

Common goals in the application of biogeochemical models of ecosystem processes are to \begin{inparaenum}[1\upshape)]
\item accurately describe the system by matching predictions with observed data \citep{Reckhow90}, and 
\item provide a means of forecasting ecosystem condition with hypothesized management or environmental scenarios \citep{Clark01}. 
\end{inparaenum}
Although these objectives are the focus of most applications, the structural components of process-based models should secondarily provide inference into which ecosystem processes and functions are driving observed changes.  This latter objective represents a more generic scientific purpose of biogeochemical models that extends beyond the applied benefits of describing and predicting change in a particular system.  Modelers often hope to identify universal principles that govern dynamics across systems and the constraint of model parameters to observations provides a means of supporting or refuting hypotheses \citep{Kirchner06}.  Extension of these principles to test the effects of structural changes and observation uncertainty on model predictions provides further information to support validation of model components. This study provided a simple approach to use the effects of parameter perturbations on model state variables to characterize identifiable parameter subsets that vary by parameter selection criteria.  By doing so, we demonstrated that small parameter subsets relative to all sensitive parameters were within the identifiability thresholds described in the literature.  The identifiable parameter subsets varied considerably between state variables and the method for parameter selection.  In general, these results provide justification for the use of explicit parameter selection heuristics that practitioners should adopt to facilitate model calibration.

@@

A general conclusion from the identifiability analyses is that only limited subsets of parameters were identifiable within the constraints of local sensitivity analyses. These results support previous studies that have suggested similarly small subsets of parameters can be identified using traditional calibration schemes \citep[e.g.,][]{Wheater86,Ye97,Omlin01}.  In addition to \ac{cgem}, these conclusions have relevance for many biogeochemical models that include numerous parameters and structural equations to characterize processes in the model domain.  A general conclusion is that perhaps a trend towards less complex models could be beneficial given that only a small subset of parameters is identifiable and that ecosystem processes may in fact be sufficiently characterized with few parameters \citep{Ye97}.  Conversely, others have argued that model complexity is not in itself a disadvantage when parsimony is not the sole arbitrator of model structure \citep{Reichert97}. Over-parameterization can be useful if processes have importance that were not evaluated during model identification.  Single objective functions that maximize model precision with identifiable parameters may also provide an incomplete characterization of model worth, which has prompted the development of probability-based models of hypoxia that explicitly include uncertainty in model components \citep[e.g.][]{Obenour15}.  Our results demonstrated that approximately 75\% of the evaluated parameters had an effect on the eight state variables, whereas \ac{cgem} includes a total of 36 variables and multiple plankton groups, not all of which have immediate concern for understanding hypoxia. The redundancies identified with the sensitivity analyses are only problematic if the primary interest is, for example, \ac{do} dynamics.  Moreover, the proposed selection heuristics provide flexibility for choosing different parameters with the assumption that those chosen depend on the research or management question.

Results from the identifiability analyses provided additional insight into the interactions of parameters in large biogeochemical models.  First, identifiability of parameter subsets was not related to the sensitivity of individual variables. As noted above, an identifiable parameter is one that has a unique effect on model predictions that cannot be compensated for or undone by changing other parameters. The magnitude of the effect of a parameter has no bearing on identifiability, which further complicates the selection of parameters for calibration.  Although identifiability is the primary limiting factor in choosing a set, the relative sensitivities are more important for the decision to include or exclude individual parameters.  Our analysis addressed this challenge by presenting multiple selection criteria for identifiable parameter sets that prioritized the most sensitive parameters during the selection process.  Similarly, identifability was not always related to the number of parameters in a set. Although the overwhelming trend was decreasing identifiability with more parameters, the unique effects of including an individual parameter with an existing set often reduced the $\gamma$ estimate. For example, \cref{fig:exclex} shows that including \Sexpr{par_txt('KGcdom_1')}, \Sexpr{par_txt('KO2_1')}, or \Sexpr{par_txt('nitmax_1')} in parameter sets more often reduced $\gamma$ relative to sets that excluded the parameters.  These examples demonstrate the complex interactions of parameter changes on variable response, highlighting the need to consider the combined and individual effects of parameters on identifiability. The selection criteria proposed in our above analyses can facilitate parameter selection and also provide diagnostic tools to identify parameters with disproportionate effects on $\gamma$.

>>>>

\begin{abstract}
\noindent Local sensitivity analyses and identifiability of parameter subsets were used to describe numerical constraints of a hypoxia model for bottom waters of the northern Gulf of Mexico.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates and less sensitive to physical or chemical parameters.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set had unique effects on the model output. Selecting parameters by decreasing sensitivity demonstrated that only eight of \Sexpr{sum(unlist(nparms))} total parameters had a sufficiently unique effect on model output for accurate calibration.  As a result, parameter selection heuristics were used to identify parameters for model calibration that depended on combined effects on output, relative sensitivity of each parameter, and ecological categories for the biogeochemical equations. The calibrated \ac{zerod} unit of the hypoxia model had improved precision if sensitive parameters were included in an identifiable subset.  Extension of results to a three-dimensional grid of the Gulf of Mexico showed that sensitive parameters for the \ac{zerod} model translated to non-trivial changes in the areal estimates of hypoxia.
\end{abstract}

@@

The development of a model represents a tradeoff between characteristics expected from the output and information represented by the structural components. An ideal model is sufficiently generalizable across systems, provides results that are precise given the inputs, and includes components that are realistic descriptions of actual processes \citep{Levins66}. Given that these characteristics cannot be simultaneously achieved, models are developed in partial dependence of reality and theoretical constructs, completely separate from both, or dependent on one or the other \citep{Morrison99,Ganju16}.  These challenges are analagous to the well-known bias-variance tradeoff in statistical models that balances the competing objectives of over- and under-fitting to an observed dataset. Process-based models are more commonly imbalanced between reality and theory, such that most are over-parameterized in an attempt to completely describe reality \citep{Denman03,Nossent12,Petrucci14}.  Quantitative limitations of over-parameterization are analagous to degrees of freedom in standard statistical models as free parameters cannot be numerically estimated when constrained to an observed dataset \citep{Kirchner06}.  More importantly, over-parameterization can limit use across systems outside of the data domain and impose uncertainty in model predictions as realistic values for every variable may not be known or inaccurately applied from existing studies \citep{Durand02,Refsgaard07,Wade08}.

@@

This study describes a parameter sensitivity and identifiability analysis of a \ac{zerod} unit of a larger spatial-temporal model of hypoxia dynamics on the \ac{lcs}.  The objectives were to provide a statistical approach that demonstrates numerical limitations of parameter sets for model calibration and provide a framework for selecting parameters within the identifiability constraints.  The specific goals were to \begin{inparaenum}[1\upshape)]
\item identify the parameters that have the greatest influence on state variables using local sensitivity analysis,
\item quantify the identifiability of subsets of the total parameter space based on sensitivity,
\item and provide a set of heuristics for choosing parameters based on sensitivity, identifiability, and parameter categories.
\end{inparaenum}
The \ac{zero} model was calibrated with selected parameter subsets to demonstrate use of the selection heuristics to improve model precision.  Sensitive parameters for the \ac{zerod} model were also varied in the larger 3-dimensional model to demonstrate scalability the results. In addition to \ac{do}, other state variables that were evaluated included ammonium, \ac{chla}, irradiance, nitrate, \ac{pom}, \ac{dom}, and phosphorus. In general, we provide empirical results to support the assumption that models are generally over-parameterized and only a finite and smaller subset of the complete parameter set can be optimized.

@@

Hypoxic events, defined  as $<$2 mg L$^{-1}$ of \ac{do} ($<$ 64 mmol m$^{-3}$), occur seasonally in bottom waters in the northern \ac{gom}.  The hypoxic area averages 15,540 km$^2$ annually (1993-2015) with minimum concentrations observed from late spring to early fall.  Seasonal variation is strongly related to carbon and nutrient export from the \ac{marb} \citep{Lohrenz08,Bianchi10}, whereas hydrologic variation, currents, and wind patterns can affect vertical salinity gradients that contribute to hypoxia formation \citep{Wiseman97,Paerl98,Obenour15}. The \ac{cgem} was developed to describe hypoxia dynamics on the \ac{lcs} and includes elements from the Navy Coastal Ocean Model \citep{Martin00} for hydrodynamics and a biogeochemical model with multiple plankton groups, water-column metabolism, and sediment diagenesis \citep{Eldridge10}.  The hydrodynamic component of \ac{cgem} provides a spatially-explicit description of hypoxia using an orthogonal grid with an approximate horizontal resolution of 1.9 km$^2$ and twenty equally-spaced vertical sigma layers on the shelf.  The biogeochemical component includes equations for 36 state variables including six phytoplankton groups (with nitrogen and phosophorus quotas for each), two zooplankton groups, nitrate, ammonium, phosphate, dissolved inorganic carbon, oxygen, silica, and multiple variables for dissolved and particulate organic matter from different sources.

@@

The sensitivity of state variables to perturbations of all relevant parameters for the 108 equations was estimated using a five minute timestep with daily output from January 1\textsuperscript{st} to December 31\textsuperscript{st}, 2006. Irrelevant parameters were removed for several reasons; parameters were not relevant for the \ac{zerod} model (i.e., hydrodynamic parameters), were considered physical constants, or had no effect given initial conditions.  Additionally, FishTank includes six phytoplankton and two zooplankton groups to add complexity in community structure and foodweb dynamics. To remove obvious redundancies, the sensitivity analyses were conducted using only one phytoplankton and one zooplankton group.  The final set that was evaluated included \Sexpr{length(p1z1pars)} parameters that were further grouped into one of six categories based on applicable biogeochemical components of the model: optics ($n = $ \Sexpr{nparms$Optics} parameters), organic matter (\Sexpr{nparms$`Organic Matter`}), phytoplankton (\Sexpr{nparms$Phytoplankton}), temperature (\Sexpr{nparms$Temperature}), and zooplankton (\Sexpr{nparms$Zooplankton}).  A full description of the model parameters is available as an appendix in \citetalias{LehrterIP}.

@@

A local sensitivity analysis was performed by evaluating the change in state variables following perturbation of each parameter from its original value \citep{Soetaert10,RDCT17}. Parameters were individually perturbed by a 50\% increase of the original values and sensitivity $S$ was estimated for each time step $i$ given a change for parameter $j$ as:

@@

Parameter selection for model calibration must consider the competing objectives of increased precision with parameter inclusion and reduced identifiability as it relates to optimization.  An additional challenge is a large number of combinations of parameter sets, which complicates selection given sensitivity differences and desired ecological categories of each parameter (e.g., parameters for a group of related structural equations could be of interest).  \cref{fig:combnex} provides a simple graphic of the unique number of combinations that are possible for different subsets of `complete' parameter sets of different sizes (i.e., $n$ choose $k$ combinations, $n!/\left(k!\left(n-k\right)!\right)$).  The number of unique combinations increases with the total parameters in the set and is also maximized for moderate selections (e.g., selecting half the total).  For example, over 10$^{14}$ combinations are possible by selecting 25 parameters from a set of 50.

A set of heuristics was developed that addresses the tradeoff in model complexity and identifiability given the challenges described above \citep[see also][]{Wagener01}.  These rulesets were developed to select parameters with preference for those with high sensitivity and identifiability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability between parameters).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences by users given desired application of the model.  For all selection heuristics, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within each of five parameter categories,
\item independent of parameter category,
\item or considering all categories equally.
\end{inparaenum}

\subsection{Model calibration and scalability}

@@

Each parameter subset was calibrated to most closely match the model output to the observed time series.  Parameter values were searched using the optim function in R that implemented the limited memory modification of the BFGS quasi-Newton method \citep{Byrd95,Nocedal06,RDCT17}.  This opimization scheme searches for values within the user-defined lower and upper bounds of each parameter.  We set the algorithm to search within +/- 50\% of the default parameter values.  While we recognize that `actual' parameter values could possibly extend beyond this range, constraining the values to the range tested herein was expected to produce output that was more interpretable within the context of the sensitivity functions defined above.  Moreover, precise limits on each parameter were less of a concern for the current analysis as our primary goal was to demonstrate differences in precision as a function of sensitivity and identifiability.  The optimal values for each parameter were identified based on convergence of the model results as a minimization of the \ac{rmse} between the observed and modelled data.

As a final analysis, effects of parameter changes on the larger 3-dimensional model were evaluated to demonstrate that results from the \ac{zerod} FishTank model are scalable.  For simplicity, only the most sensitive parameter in each category was selected.  The \ac{cgem} model was run for one year using the default parameter conditions and by increasing each of the selected parameters by 50\%.  This produced six spatially-referenced time series (one default, five for each parameter change) from which changes in the daily hypoxic area (number of cells $<$ 64 mmol \ac{do}) were estimated.

@@

The identifiability analyses suggested that many parameter subsets exceeded the thresholds of $\gamma = 10, 15$.  Parameter identifiability for \ac{do} decreased (increasing $\gamma$) at different rates with increasing size of parameter subsets depending on the parameter category or the number of top parameters that were selected (\cref{fig:identplo}). By category, identifiability was lowest for all combinations of parameter subsets in the phytoplankton (\Sexpr{catpersO2$Phytoplankton[1]}\% of subsets less than $\gamma = 15$, \Sexpr{catpersO2$Phytoplankton[2]}\% less than $\gamma = 10$) and zooplankton categories (\Sexpr{catpersO2$Zooplankton[1]}\% less than $\gamma = 15$, \Sexpr{catpersO2$Zooplankton[2]}\% less than $\gamma = 10$), whereas all combinations were identifiable for optics (\Sexpr{catpersO2$Optics[1]}\% less than $\gamma = 15, 10$) and a majority identifiable for organic matter (\Sexpr{catpersO2$`Organic Matter`[1]}\% less than $\gamma = 15$, \Sexpr{catpersO2$`Organic Matter`[2]}\% less than $\gamma = 10$). Identifiability for parameters in the temperature category was not evaluated because \ac{do} was sensitive to only one parameter (i.e., $\gamma = 1$). Parameter combinations for choosing from the top, top two, top three, and top four parameters in each category together had decreasing identifiability with the increasing size of the selection pool (e.g., top one versus top four parameters, \cref{fig:identplo}).  The percentage of parameter subsets that were below the acceptable thresholds for identifiability was \Sexpr{toppersO2$top1[1]}\% less than $\gamma = 15, 10$ for the top parameter in each category, \Sexpr{toppersO2$top2[1]}\% and \Sexpr{toppersO2$top2[2]}\% for the top two, \Sexpr{toppersO2$top3[1]}\% and \Sexpr{toppersO2$top3[2]}\% for the top three, and \Sexpr{toppersO2$top4[1]}\% and \Sexpr{toppersO2$top4[2]}\% for the top four.  Results for the remaining state variables had similar patterns in identifiability with increasing size of parameter subsets and selection categories, although differences in identifiability between state variables was observed (\cref{fig:identploall}).  Most notably, nitrate was consistently the least identifiable variable (highest overall $\gamma$), whereas \ac{do} was most identifiable.

@@

\subsection{Model calibration and scalability}

@@

Calibration of parameter values for each of the seven selection heuristics (\cref{tab:heurist1,tab:heurist2,tab:heurist3})  produced model output that varied in precision and computational effort required to identify optimal values (\cref{tab:calib}).  As expected, the number of required iterations of the optimization algorithm to minimize \ac{rmse} of observed \ac{do} and model output varied in proportion to the number of parameters in a subset.  For example, only \Sexpr{toeng(temp$its)[[1]]} iterations were required to optimize the single parameter selected for the temperature category, whereas \Sexpr{top$its} iterations were required for the \Sexpr{toeng(top$nprm)[[1]]} parameters selected independent of category.  A slight positive association between collinearity ($\gamma$) and number of iterations was also observed, such that computational effort  to identify optimal values for a parameter set increased with lower identifiability (higher $\gamma$).  Of the seven parameter subsets, selection independent of parameter category and equally within each category produced the largest reduction in \ac{rmse} values from the default output (starting \ac{rmse} \Sexpr{strt}, reduction to \Sexpr{scinot(top$rmse, digits = 3)}, \Sexpr{scinot(cattop$rmse, digits = 3)}, respectively), whereas the optimal values for parameters in the Optics category had the smallest reduction (\Sexpr{scinot(optics$rmse, digits = 3)}).  Lower \ac{rmse} was associated with higher average sensitivity of a parameter subset.  Parameters selected only in the optics category had the highest \ac{rmse} and lowest average L1 (\Sexpr{scinot(optics$avee)}, \Sexpr{toeng(optics$nprm)} parameters), whereas parameters selected independent of category had the lowest \ac{rmse} and highest average L1 (\Sexpr{scinot(top$avee)}, \Sexpr{toeng(top$nprm)} parameters, \cref{tab:calib}).

@@

Common goals in the application of biogeochemical models are to \begin{inparaenum}[1\upshape)]
\item accurately describe the system by matching predictions with observed data \citep{Reckhow90}, and 
\item provide a means of forecasting ecosystem condition with hypothesized management or environmental scenarios \citep{Clark01}. 
\end{inparaenum}
Although these objectives are the focus of most applications, the structural components and associated parameters of process-based models should secondarily provide insight of ecosystem processes that are driving observed changes.  This latter objective represents a more generic scientific purpose of modelling that extends beyond the benefits of describing and predicting change.  Modelers aim to identify universal principles that govern dynamics across systems and the constraint of model parameters to observations provides a means of supporting or refuting hypotheses \citep{Hilborn97,Kirchner06}.  Accordingly, this study provided a simple approach to use the effects of parameter perturbations on state variables to characterize identifiable parameter subsets that varied by selection criteria.  By doing so, we provided a simple and preliminary approach towards improved model refinement that begins to address more general purposes of model applications. Our specific results demonstrated that small parameter subsets relative to all sensitive parameters were within the identifiability thresholds described in the literature.  The identifiable parameter subsets varied considerably between state variables and the method for parameter selection.  Use of identifiable parameter subsets with sensitive parameters improved model precision and results for the smaller model were scalable to the 3-dimensional model.  In general, these results provide justification for the use of explicit parameter selection heuristics to improve model precision.

@@

A general conclusion from the identifiability analyses is that only limited subsets of parameters were identifiable within the constraints of local sensitivity analyses. These results support previous studies that have suggested similarly small subsets of parameters can be identified using traditional calibration schemes \citep[e.g.,][]{Wheater86,Ye97,Omlin01}.  In addition to \ac{cgem}, these conclusions have relevance for other biogeochemical models that include numerous parameters and structural equations to characterize processes in the model domain.  A general conclusion is that less complex models could potentially be beneficial given that only a small subset of parameters is identifiable and that ecosystem processes may in fact be sufficiently characterized with few parameters \citep{Ye97}.  Conversely, others have argued that model complexity is not in itself a disadvantage when parsimony is not the sole arbitrator of model structure \citep{Reichert97}. Over-parameterization can be useful if processes have importance that were not evaluated during model identification.  Single objective functions that maximize model precision with identifiable parameters may also provide an incomplete characterization of model worth, which has prompted the development of probability-based models of hypoxia that explicitly include uncertainty in model components \citep[e.g.][]{Obenour15}.  Our results demonstrated that approximately 75\% of the evaluated parameters had an effect on the eight state variables, whereas \ac{cgem} includes a total of 36 variables and multiple plankton groups, not all of which have immediate concern for understanding hypoxia. The redundancies identified with the sensitivity analyses are only problematic if the primary interest is, for example, \ac{do} dynamics.  Moreover, the proposed selection heuristics provide flexibility for choosing different parameters with the assumption that those chosen depend on the research or management question.

Results from the identifiability analyses provided additional insight into the interactions of parameters in large biogeochemical models.  First, identifiability of parameter subsets was not related to the sensitivity of individual variables. As noted above, an identifiable parameter is one that has a unique effect on model predictions that cannot be compensated for or undone by changing other parameters. The magnitude of the effect of a parameter has no bearing on identifiability, which further complicates the selection of parameters for calibration.  Although identifiability is the primary limiting factor in choosing a set, the relative sensitivities are more important for the decision to include or exclude individual parameters.  For example, \cref{tab:calib,fig:areachg} demonstrated that parameter sets with sensitive parameters had improved model precision and that these parameters had affects that translated to the larger model. Our analysis addressed this challenge by presenting multiple selection criteria for identifiable parameter sets that prioritized the most sensitive parameters during the selection process.  Similarly, identifiability was not always related to the number of parameters in a set. Although the overwhelming trend was decreasing identifiability with more parameters, the unique effects of including an individual parameter with an existing set often reduced the $\gamma$ estimate. For example, \cref{fig:exclex} showed that including \Sexpr{par_txt('KGcdom_1')}, \Sexpr{par_txt('KO2_1')}, or \Sexpr{par_txt('nitmax_1')} in parameter sets more often reduced $\gamma$ relative to sets that excluded the parameters.  These examples demonstrate the complex interactions of parameter changes on variable response, highlighting the need to consider the combined and individual effects of parameters on identifiability. The selection criteria proposed in our above analyses can facilitate parameter selection and also provide diagnostic tools to identify parameters with disproportionate effects on $\gamma$.
